---
title: "Faye Harwell- AN597- Module 8"
author: "Faye Harwell"
date: "September 26, 2017"
output: html_document
---
# Plotting a Histogram using Manipulate()

### I installed manipulate prior to running this output

### Within hist(), you can determine where you want breaks

### Very cool- in manipulate, you can create a slider, where you can manipulate n. Look down at the code where n = slider (0,10000,initial=100,step=100)

```{r}
library(manipulate)
outcomes <- c(1, 2, 3, 4, 5, 6)
manipulate(hist(sample(outcomes, n, replace = TRUE), breaks = c(0.5, 1.5, 2.5, 
    3.5, 4.5, 5.5, 6.5), probability = TRUE, main = paste("Histogram of Outcomes of ", 
    n, " Die Rolls", sep = ""), xlab = "roll", ylab = "probability"), n = slider(0, 
    10000, initial = 100, step = 100))
```

### CHALLENGE: Write a function to simulate rolling a die where you pass the number of rolls as an argument. Then, use your function to simulate rolling two dice 1000 times and take the sum of the rolls. Plot a histogram of those results.

```{r}
nrolls <- 1000
roll <- function(x) {
    sample(1:6, x, replace = TRUE)
}
two_dice <- roll(nrolls) + roll(nrolls)
hist(two_dice, breaks = c(1.5:12.5), probability = TRUE, main = "Rolling Two Dice", 
    xlab = "sum of rolls", ylab = "probability")
```

# Heads or Tails? Flipping a Coin

```{r}
outcomes <- c("heads", "tails")
prob <- c(1/2, 1/2)
barplot(prob, ylim = c(0, 0.6), names.arg = outcomes, space = 0.1, xlab = "outcome", 
    ylab = "Pr(X = outcome)", main = "Probability Mass Function")
```

```{r}
cumprob <- cumsum(prob)
barplot(cumprob, names.arg = outcomes, space = 0.1, xlab = "outcome", ylab = "Cumulative Pr(X)", main = "Cumulative Probability")
```

# Rolling a Fair Die... do we really need to specify that a die is fair?

```{r}
outcomes <- c(1, 2, 3, 4, 5, 6)
prob <- c(1/6, 1/6, 1/6, 1/6, 1/6, 1/6)
barplot(prob, ylim = c(0, 0.5), names.arg = outcomes, space = 0.1, xlab = "outcome", 
    ylab = "Pr(X = outcome)", main = "Probability Mass Function")
```

#Cumulative Probability... adding rolls?

```{r}
cumprob <- cumsum(prob)
barplot(cumprob, names.arg = outcomes, space = 0.1, xlab = "outcome", ylab = "Cumulative Pr(X)",  main = "Cumulative Probability")
```

# Beta Distributions- you can manipulate the values of alpha and beta

```{r}
library(ggplot2)
a <- 2
b <- 1
K <- 2
x <- seq(from = 0, to = 1, by = 0.025)
fx <- K * x^(a - 1) * (1 - x)^(b - 1)
lower_x <- seq(from = -0.25, to = 0, by = 0.025)  # add some values of x less than zero
upper_x <- seq(from = 1, to = 1.25, by = 0.025)  # add some values of x greater than one
lower_fx <- rep(0, 11)  # add fx=0 values to x<0
upper_fx <- rep(0, 11)  # add fx=0 values to x>1
x <- c(lower_x, x, upper_x)  # paste xs together
fx <- c(lower_fx, fx, upper_fx)  # paste fxs together
d <- as.data.frame(cbind(x, fx))
p <- ggplot(data = d, aes(x = x, y = fx)) + xlab("x") + ylab("f(x)") + geom_line()
p
```

```{r}
library(manipulate)
manipulate(ggplot(data = d, aes(x = x, y = fx)) + xlab("x") + ylab("f(x)") + 
    geom_line() + geom_polygon(data = data.frame(xvals = c(0, n, n, 0), fxvals = c(0, 
    K * n^(a - 1) * (1 - n)^(b - 1), 0, 0)), aes(x = xvals, y = fxvals)) + ggtitle(paste("Area Under Function = ", 
    0.5 * n * K * n^(a - 1) * (1 - n)^(b - 1), sep = " ")), n = slider(0, 1, 
    initial = 0.5, step = 0.01))
```
```{r}
x <- seq(from = 0, to = 1, by = 0.005)
prob <- 0.5 * x * K * x^(a - 1) * (1 - x)^(b - 1)
barplot(prob, names.arg = x, space = 0, main = "Cumulative Probability", xlab = "x",  ylab = "Pr(X ≤ x)")
```
 ### The built in R function for the Beta Distribution, pbeta(), can give us the cumulative probability directly, if we specify the values of αα = 2 and ββ = 1.
 
```{r}
pbeta(0.75, 2, 1) 
```

# Playing around with probabilities

```{r}
n <- 6  # number of trials
k <- 6  # number of successes
p <- 1/6
prob <- (factorial(n)/(factorial(k) * factorial(n - k))) * (p^k) * (1 - p)^(n - k)
prob 
```

```{r}
k <- 3  # number of successes
prob <- (factorial(n)/(factorial(k) * factorial(n - k))) * (p^k) * (1 - p)^(n - k)
prob
```

### You could also do...

```{r}
dbinom(x = k, size = n, prob = p)
```

## To get the cumulative probability... do this...

```{r}
probset <- dbinom(x = 0:6, size = 6, prob = 1/6)  # x is number of successes, size is number of trials
barplot(probset, names.arg = 0:6, space = 0, xlab = "outcome", ylab = "Pr(X = outcome)",  main = "Probability Mass Function")
```

```{r}
cumprob = cumsum(probset)
barplot(cumprob, names.arg = 0:6, space = 0.1, xlab = "outcome", ylab = "Cumulative Pr(X)", main = "Cumulative Probability")
```

```{r}
sum(probset)
```

### The chance of observing exactly 3 rolls of 1 is...

```{r}
dbinom(x = 3, size = 6, prob = 1/6)
```

## To figure out the distribution of your data, you can graph it and sees if it falls into any of the categories (binomial, poisson, bernoulli, etc.)

# Poisson Distribution- often used to model open ended counts of independently occuring events, for example the number of cars that pass a traffic intersection over a given interval of time or the number of times a monkey scratches itself during a given observation interval. The probability mass function for the Poisson distribution is described by a single parameter, λ, where λ can be interpreted as the mean number of occurrences of the event in the given interval.

## Using R to look at the probability mass functions for different values of λ

```{r}
x <- 0:10
l = 3.5
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = "Probability Mass Function")
```

```{r}
x <- 0:20
l = 10
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = "Probability Mass Function")
```

```{r}
x <- 0:50
l = 20
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = "Probability Mass Function")
```

### As we did for other distributions, we can also use the built in probability function for the Poisson distribution, ppois(), to return the value of the cumulative distribution function, i.e., the probability of observing up to and including a specific number of events in the given interval.

```{r}
x <- 0:10
l <- 3.5
barplot(ppois(q = x, lambda = l), ylim = 0:1, space = 0, names.arg = x, xlab = "x", 
    ylab = "Pr(X ≤ x)", main = "Cumulative Probability")
```

```{r}
x <- 0:20
l <- 10
barplot(ppois(q = x, lambda = l), ylim = 0:1, space = 0, names.arg = x, xlab = "x", 
    ylab = "Pr(X ≤ x)", main = "Cumulative Probability")
```

```{r}
x <- 0:50
l <- 20
barplot(ppois(q = x, lambda = l), ylim = 0:1, space = 0, names.arg = x, xlab = "x", 
    ylab = "Pr(X ≤ x)", main = "Cumulative Probability")
```